
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{ACKNOWLEDGMENTS}
%\addcontentsline{toc}{chapter}{ACKNOWLEDGMENTS}
\vspace{7mm}

\begin{doublespace}

When I began to write this section, I quickly realized that to truly
acknowledge everyone that merits acknowledgement I would have to recount
my entire life story;
where I am now is a complicated function
of every decision that I, and everyone else, has made.
Since my memory isn't good enough to recall every decision, I can only do my
best to thank all of those that have contributed to this dissertation,
both tangibly and intangibly.

Firstly, I want to acknowledge that it is not always the case that everyone
a student encounters is supportive and has the student's best interests
in mind and at heart.
I have been extremely fortunate that in my case, that has been true.
I could not have asked for better people to be guiding me in my career.

In roughly chronological order, I would first like to thank Mr. Kapp,
my first college-level physics teacher, who
helped to keep my physics knowledge grounded in practicality.
I would next like to thank Mr. Quail, my calculus II teacher,
who really let me know that I had
a natural aptitude for math, even at the college level.
Once I graduated from Washtenaw Community College and transferred to the
University of Michigan, professor Keren Sharon put me on the research track
by asking me if I would like to work with her.
I would not have had the confidence to ask a professor such a question,
so that was a pivotal moment for my career.
Keren went on to be my undergraduate advisor, and I want to thank her for
introducing me to the field of astronomy research and for
helping me navigate my career as a young academic.
When I didn't get into any graduate schools, professor Nuria Calvet went out
of her way to mention me to Keivan Stassun, a professor at
Vanderbilt University,
who was a co-founder of the Fisk-Vanderbilt Master's-to-PhD Bridge Program.
A second pivotal moment happened when I applied and was accepted into the
Bridge program, and thanks to that goes to Nuria (and a little to me,
for being so great that I was accepted).
The Bridge program, as a whole, is an immensely supportive environment
and I am very fortunate to have been a part of it.
Once in the Bridge program,
I met Kelly Holley--Bockelmann, a professor who would become the
co-chair of both my Master's and Ph.D. dissertation committees.
Kelly introduced me to professor Anthony Mezzacappa,
who would also serve on both of my
committees, and professor Eirik Endeve,
who would be co-chair of both of my committees.
I would also like to thank Dr. Lauren Campbell for all she has done for
me, in particular in her role as the director of the Bridge Program.

Of course, a dissertation requires much more than academic guidance.
I also thank my wife, Kate, for her unwavering support throughout this journey.
It is unfortunately not the case that one's spouse is always so
supportive, and I am very lucky that Kate has been.
I also thank my parents, Sandy and Steve, for their support and encouragement.
My life, like most people's, has been a bit of an emotional rollercoaster,
and they have always been there and have been supportive of my decisions.
I also want to thank my brother, Cal, to whom I have looked up since I was
very young and continue to do so today.
I thank my grandparents, Jack and Jean, who always knew the value
of education and who helped me begin my academic journey with a college fund.
I know they would be beyond proud to see how far I've come.

In addition to academic and family support, my journey has been greatly
helped by my friends and peers.
Moving to a new state is tough, and
getting through the graduate classes was much easier because of my
bridge cohort, Dax, Antonio, and Natasha, as well as other members of the
Vanderbilt astronomy group, Savannah, John, Karl, George, Laura, and Victor.
They have all helped me get through this, and I am grateful to them all.
Thanks also to my childhood friends from Michigan,
Jeff, Evan, Travis, Kris, and Scott.
They've always been there for me and always made coming home extra special.

Lastly, I thank my pets, Lily, Daisy, Bo, and Penny.
They all gave and continue to give so much joy and fun and enrich
my life tremendously.

My sincerest apologies to anyone I left out, but please know that I appreciate
you.

\end{doublespace}
